BarnebyLives: an R package to create herbarium specimen labels and clean spreadsheets

### 2025-04-23 manuscript/BarnebyLives An R package to create and qc herbarium labels.rmd
back to manuscript!

data <- read.csv('data/test_data.csv', na.strings = "") %>% 
  drop_na(c('Longitude', 'Latitude', 'Date_digital')) %>% 
  unite(col = 'Scientific_name', c(Binomial, Infrarank, Infraspecies), na.rm=TRUE, sep = " ", remove = F)

@@

#time_directions_grabber <- system.time({ # write directions to sites
#  SoS_gkey = Sys.getenv("Sos_gkey")
#  data <- directions_grabber(data, api_key = SoS_gkey)
#})

#saveRDS(time_directions_grabber, file = 'data/processed/time_directions_grabber')
saveRDS(data, file = 'data/processed/data_w_Google_Maps')

@@

# how many of my genera were mis-spelled??
genera_mispelled <- data_p %>% 
  sf::st_drop_geometry() %>% 
  filter(POW_Genus == 'NOT FOUND') %>% 
  select(Collection_number, Scientific_name, Genus, POW_Genus)

# the mismatch between POW genus and not
genus_mismatch <- data_p %>% 
  sf::st_drop_geometry() %>% 
  filter(POW_Genus != 'NOT FOUND') %>% 
  filter(POW_Genus != Genus) %>% 
  select(Collection_number, Full_name, Genus, POW_Genus) %>% 
  mutate(Situation = 
    case_when(
      Genus %in% c('Glandularia', 'Oenothera', 'Chamerion') ~ 'Preferred',
      # check Pleuraphis, i am confused on this still....
      Genus %in% c('Cryptantha', 'Ivesia', 'Lotus', 'Polygala', 'Minuartia', 'Pleuraphis',
                    'Pseudostellaria', 'Achnatherum', 'Scirpus', 'Peritoma', 'Arenaria', 
                    'Spartina', 'Polygonum', 'Atriplex', 'Cymopterus', 'Trisetum', 'Bahia', 
                   'Oryzopsis', 'Potentilla', 'Pascopyrum'
                    ) ~ 'Outdated',
      Genus %in% c('Vaccinium', 'Eleocharis') ~ 'Internal_Error',
      .default = 'Typo'
    ),
    Situation = if_else(Full_name == 'Oenothera caespitosa Nutt.', 'Internal_Error', Situation))

@@

epithet_diff <- data_p %>% 
  sf::st_drop_geometry() %>% 
  filter(Epithet != POW_Epithet, !Full_name %in% genus_mismatch$Full_name) %>% 
  select(Collection_number, Full_name, Epithet, POW_Genus, POW_Epithet, POW_Infraspecies) %>% 
  distinct(Full_name, .keep_all = T) %>%  # do we record repeat collections ?
  mutate(Situation = 
           case_when(
             Collection_number %in% c(
               '1217', '1251', '1283', '1311', '1337', '1391', '1450', '1576', 
               '1594', '1745', '2043', '2086', '2108', '2109', '2120', '2127', '2130',
               '2153', '2199', '2227', '2242', '2256', '2289', '2297', '2343', '2370', '2379', 
               '2393', '2415', '2480', '2511', '2561', '2573', '2648', '2660', 
               '2687', '2690', '2710', '2715', '2727', '2750', '2767', '2818')
             ~ 'Spelling',
             Collection_number %in% c(
               '1588', '1619', '2445', '2777', '1337', '2412', '2487', '2748', '2652',
               '2805', '2194', '2040', '2039', '2403', '2328', '2076') 
             ~ 'Outdated', 
             Collection_number %in% c('2644') ~ 'Incorrect',
             Collection_number %in% c(
               '1207', '2326', '2038', '1424', '2618', '2622', '2665', '2711', '2713') 
             ~ 'Internal_Error',
             Collection_number %in% c('2110', '2540') ~ 'Preferred'
           )
  )

@@

Of the `r nrow(data)` records analysed (`r n_spp` distinct species) BarnebyLives flagged `r nrow(epithet_diff)` records, and detected `r sum(epithet_diff$Situation=='Spelling')`  instances of misspelled epithets (`r sum(epithet_diff$Situation=='Spelling')` unique species). 
Of the `r nrow(data)` species analysed (`r n_spp` distinct species) BarnebyLives flagged `r nrow(epithet_diff)` records, and detected `r sum(epithet_diff$Situation=='Spelling')`  instances of misspelled epithets (`r sum(epithet_diff$Situation=='Spelling')` unique species). 
In `r sum(epithet_diff$Situation=='Outdated')` of these instances the user supplied an outdated name (`r nrow(unique(epithet_diff[epithet_diff$Situation=='Outdated', 'Full_name']))` unique species). 
It also flagged `r sum(epithet_diff$Situation=='Preferred', na.rm = T)` records where the author follows an alternative taxonomy (`r nrow(unique(epithet_diff[epithet_diff$Situation=='Preferred', 'Full_name']))` unique species), and flagged `r sum(epithet_diff$Situation=='Internal_Error', na.rm = T)` records in error. 
The final record was an egregious error where the order of the specific epithet and the genus name.

>>>>

#subs_batch <- c(1384, 1388, 1394, 1396, 1397, 1400, 1435, 1438, 2161, 2187, 2268, 
#                2301, 2302, 2577, 2746, 2785, 1460, 1461, 1379, 1381)

data <- read.csv('data/test_data.csv', na.strings = "") %>% 
  drop_na(c('Longitude', 'Latitude', 'Date_digital')) %>% 
  unite(col = 'Scientific_name', 
        c(Binomial, Infrarank, Infraspecies), na.rm=TRUE, sep = " ", remove = F) %>% 
  filter(Scientific_name != '')#, Collection_number %in% subs_batch)#, Collection_number %in% subs_batch)

@@

time_directions_grabber <- system.time({ # write directions to sites
  SoS_gkey = Sys.getenv("Sos_gkey")
  data <- directions_grabber(data, api_key = SoS_gkey)
})

saveRDS(time_directions_grabber, file = 'data/processed/time_directions_grabber')
saveRDS(data, file = 'data/processed/data_w_Google_Maps')

@@

# how many of my genera were mis-spelled??

## HEY REED SPLIT OUT GENERA FOR THIS !!!!! AGAIN .... 
data_p <- split_scientificName(data_p, sciName_col = 'Scientific_name')

genera_mispelled <- data_p %>% 
  sf::st_drop_geometry() %>% 
  filter(POW_Genus == 'NOT FOUND') %>% 
  select(Collection_number, Scientific_name, Genus, POW_Genus)

# the mismatch between POW genus and not
genus_mismatch <- data_p %>% 
  sf::st_drop_geometry() %>% 
  filter(POW_Genus != 'NOT FOUND') %>% 
  filter(POW_Genus != Genus) %>% 
  select(Collection_number, Scientific_name, Genus, POW_Genus) %>% 
  mutate(Situation = 
    case_when(
      Genus %in% c('Glandularia', 'Oenothera', 'Chamerion') ~ 'Preferred',
      # check Pleuraphis, i am confused on this still....
      Genus %in% c('Cryptantha', 'Ivesia', 'Lotus', 'Polygala', 'Minuartia', 'Pleuraphis',
                    'Pseudostellaria', 'Achnatherum', 'Scirpus', 'Peritoma', 'Arenaria', 
                    'Spartina', 'Polygonum', 'Atriplex', 'Cymopterus', 'Trisetum', 'Bahia', 
                   'Oryzopsis', 'Potentilla', 'Pascopyrum'
                    ) ~ 'Outdated',
      Genus %in% c('Vaccinium', 'Eleocharis') ~ 'Internal_Error',
      .default = 'Typo'
    ),
    Situation = if_else(Scientific_name == 'Oenothera caespitosa Nutt.', 'Internal_Error', Situation))

@@

epithet_diff <- data_p %>% 
  sf::st_drop_geometry() %>% 
  filter(Epithet != POW_Epithet, !Scientific_name %in% genus_mismatch$Scientific_name) %>% 
  select(Collection_number, Scientific_name, Epithet, POW_Genus, POW_Epithet, POW_Infraspecies) %>% 
  distinct(Scientific_name, .keep_all = T) %>%  # do we record repeat collections ?
  mutate(Situation = 
           case_when(
             Collection_number %in% c(
               '1217', '1251', '1283', '1311', '1337', '1391', '1415', '1421', '1435', '1450', '1576', 
               '1594', '1624', '1627', '1745', '2043', '2086', '2108', '2109', '2120', '2127', '2130', '2135', 
               '2153', '2199', '2227', '2242', '2256', '2289', '2297', '2315', '2343', '2370', '2379', 
               '2393', '2415', '2480', '2511', '2561', '2573', '2648', '2660', 
               '2687', '2690', '2710', '2715', '2727', '2750', '2767', '2818')
             ~ 'Spelling',
             Collection_number %in% c(
               '1556', '1588', '1619', '2445', '2777', '1337', '2004', '2412', '2487', '2748', '2652',
               '2805', '2194', '2040', '2039', '2403', '2328', '2076') 
             ~ 'Outdated', 
             Collection_number %in% c('2644') ~ 'Incorrect',
             Collection_number %in% c(
               '1207', '2326', '2038', '1424', '2618', '2622', '2665', '2711', '2713') 
             ~ 'Internal_Error',
             Collection_number %in% c('2110', '2540') ~ 'Preferred'
           )
  )

@@

Of the `r nrow(data)` records analysed (`r n_spp` distinct species) BarnebyLives flagged `r nrow(epithet_diff)` records, and detected `r sum(epithet_diff$Situation=='Spelling')`  instances of misspelled epithets (`r sum(epithet_diff$Situation=='Spelling')` unique species). 
Of the `r nrow(data)` species analysed (`r n_spp` distinct species) BarnebyLives flagged `r nrow(epithet_diff)` records, and detected `r sum(epithet_diff$Situation=='Spelling')`  instances of misspelled epithets (`r sum(epithet_diff$Situation=='Spelling')` unique species). 
In `r sum(epithet_diff$Situation=='Outdated')` of these instances the user supplied an outdated name (`r nrow(unique(epithet_diff[epithet_diff$Situation=='Outdated', 'Scientific_name']))` unique species). 
It also flagged `r sum(epithet_diff$Situation=='Preferred', na.rm = T)` records where the author follows an alternative taxonomy (`r nrow(unique(epithet_diff[epithet_diff$Situation=='Preferred', 'Scientific_name']))` unique species), and flagged `r sum(epithet_diff$Situation=='Internal_Error', na.rm = T)` records in error. 
The final record was an egregious error where the order of the specific epithet and the genus name.
